### 2020-04-21 dataprep.tex
done for tonight

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Data Collection and Preprocessing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Collection and Preprocessing}
\label{section:dataprep}
Data on GPU life times is constructed from two sources: inventory runs
and failure event records. Two types of failure events were collected:
Double Bit Error (DBE) and Off the Bus (OTB). DBE was found to be the
``signature'' error of eventual GPU failures and a trigger for their
replacement. \fix{RA: Should we point out that this replacement was only done later on?}

An inventory of GPU serial numbers and their locations is recorded
each time the system boots. \fix{Don, you mentioned that this was more
regular later - daily?} This typically occurs once every few days but
can be as frequent as a few times per day. An inventory at boot time
typically occurs over a span of \fix{??} minutes. A separate file is
produced at each boot time.

>>>>

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Data Collection and Preprocessing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Collection and Preprocessing}
\label{section:dataprep}
Data on GPU life times is constructed from two sources: inventory runs
and failure event records. Two types of failure events were collected:
Double Bit Error (DBE) and Off the Bus (OTB). Both errors were found
to be the ``signature'' event of GPU failures and a trigger for their
replacement.

An inventory of GPU serial numbers and their locations is recorded
each time the system boots. This typically occurs once every few days
but can be as frequent as a few times per day. An inventory at boot
time typically occurs over a span of \fix{how many ??} minutes. A
separate file is produced at each boot time.

### 2020-04-21 intro.tex
done for tonight

The failures were traced to a failing resistor on the circuitboard
(not the GPU iself). While the 2016 resistor failures were a surprise, there is a known
phenomenon~\cite{referenceneeded} that describes a two phase process
that leads to failure. It also explains why there was a very stable
period preceding the failures. There is specific decay associated
with the electronic components where, as these products age, they will
eventually fail thresholds for guard band margin and other things. In
a lot of cases, they will just fail as well. \fix{Give a reference and
  expand?}

Our analysis focuses on GPUs that were installed in the second rework
cycle and later, essentialy units installed near the beginning of 2014
and later. \fix{Will we (are we required) make the data publicly
  available?} The data used in this study, representing over 100,000
years of operation may be the largest publicly available GPU
reliability data set.

>>>>

The failures were traced to a failing resistor on the circuitboard
(not the GPU chip iself). Throughout this paper, when we refer to
replacing the GPU, we refer to the entire circuitboard along with its
GPU chip. While the 2016 resistor failures were a surprise, there is a
known phenomenon~\cite{referenceneeded} that describes a two phase
process that leads to failure. It also explains why there was a very
stable period preceding the failures. There is specific decay
associated with the electronic components where, as these products
age, they will eventually fail thresholds for guard band margin and
other things. In a lot of cases, they will just fail as
well. \fix{Give a reference and expand?}

Our analysis focuses on GPUs that were installed in the second rework
cycle and later, essentialy units installed near the beginning of 2014
and later. We will make the data as well as our analysis codes
producing the graphics available in a public repository (see
\fix{reference?}).  The data used in this study, while only a 7 MB
file, represents over 100,000 years of operation and may be the
largest publicly available GPU reliability data set.

### 2020-04-21 survival.tex
done for tonight

If $T$ is the random variable of a GPU fail time, then its cumulative
distribution function $F(t) = Pr\{T < t\}$ gives the probability that
a GPU fails by duration $t$. The survival function is its complement
\begin{displaymath}
  S(t) = Pr\{T \geq t\} = 1 - F(t).
\end{displaymath}
It is the probability of being operational at duration $t$.  We use
the R packages \pkg{survival} and \pkg{survminer} for the KM analysis,
which is reported in Figure~\ref{fig:km-all-cage-node}. Within each
{\em batch}, separate survival curves are computed for each {\em cage}
by {\em node} combination.

Another standard survival data methodology is a Cox proportional
hazards (CPH) regression analys, which can include covariates and
estimate relative risk based on the covariates
\cite{Cox1972,Harrell2015}. The CPH regression function takes the form
\begin{displaymath}
  h(t) = h_0(t)e^{b_1 x_1 + b_2 x_2 \ldots b_k x_k},
\end{displaymath}
where $x_i$ are covariates, $h_0(t)$ is the {\em baseline hazard}, and
the $b_i$ are coefficients that measure the impact of the
covariates. The quantity $e^{b_i}$ is the hazard ratio for covariate
$i$. Its interpretation

That is, a base hazard rate multiplied by a function of
covariates. The key assumption in the CPH model is that the "hazards"
are multipliers on the "baseline hazard" but the estimate of the
baseline hazard is nonparametric, meaning it makes no specific
distributional assumption and just learns from the data (unlike
analyses that assume for example a Weibul or an exponential
model). The multiplicative assumption is that the hazard curves do not
cross and are multiples of each other. The multiplier is the *hazard
coefficient*. For example, if the baseline is *node0* and the *node2*
hazard is 2, then *node2* sees twice as many failures as *node0* on
average.

For our comparison of hazards, we use as *covariates* all the unpacked
location variables, indicators whether the GPU has been moved, and our
notion of old and new GPUs based on first install date. Since many
GPUs experience time at more than one location, we use the longest
period location in this analysis.

First, we do a univariate analysis for each covariate separately and
then we look at a multivariate view of this. Having multiple
covariates is called "multiple regression" as opposed to "multivariate
regression" reserved for cases with several response variables. The
univariate view is comparable to what would one see in a simple time
between failures analysis for a given population split. It can be
affected by sampling differences for other covariates. For example, if
the strongest effect is seen in a variable not used in a split and its
values are unevenly distributed across the split, its effect can leak
into the split. This is why we also do a multivariate analysis, where
the effect of each covariate is adjusted for all the other
covariates.

\begin{figure}
  \centering
  \includegraphics[width=0.8\columnwidth]{figs/cox_o001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt old}
    batch.}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=0.8\columnwidth]{figs/cox_n001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt new}
    batch.}
\end{figure}

Higher numbered columns appear to have a higher (up to about 2x)
failure rate. There actually seems to be a hint of three groups that
have similar rates: 0-6 baseline, 7-16 are 1.5x baseline, and 17-24
are 2x baseline.

Higher numbered rows tend to have a lower (up to about 0.5x) failure rate.

This seems surprisingly strong! Higher numbered cages have failure
rates 5x and 16x higher than cage = 0. In fact, this seems so strong
that its effect could be leaking into all other comparisons under
uneven sampling!

Not much difference between slots. They are all close to the baseline
slot0.

Interestingly, nodes 2 and 3 have half the failure rate of nodes 0 and
1. Is this related to the cage variable or is there a real location
effect?

Now the multivariate Cox regression analysis, where each factor is
adjusted for all the other factors. This can fix some uneven sampling
and multivariate dependence.

The col groups persist and are even stronger. The row effect is also
stronger. The cage effect is out of this world! slot is still without
any effect. node effect also persists and is a bit stronger.

Let’s separate the “old” and “new” batches of GPUs and run separate
Cox models.

>>>>

If $T$ is the random variable of a GPU fail time, then its cumulative
distribution function $F(t) = Pr\{T < t\}$ gives the probability that
a GPU fails by duration $t$. The survival function is its complement
\begin{displaymath}
  S(t) = Pr\{T \geq t\} = 1 - F(t).
\end{displaymath}
It is the probability of being operational at duration $t$.  We use
the R packages \pkg{survival} and \pkg{survminer} for the KM analysis,
which is reported in Figure~\ref{fig:km-all-cage-node}. Within each
{\em batch}, separate survival curves are computed for each {\em cage}
by {\em node} combination. Along with the survival curve estimate,
this analysis provides 95\% confidence region for survival probability
shaded around the curves.

It appears that transport of cooling air provides a complete
explanation for relative differences in cage and node survival rates
in the old batch.  We reach this conclusion with reference to
Figure~\ref{fig:layout} and relative positions of cages and nodes
within a cabinet. Both cage and node differences in survival
probabilities can be explained by an inverse relationship with the
distance to the bottom of the cabinet, where cooling air is forced
through the cabinet to the top. The survival curves can be ordered
({\em cage 0, cage 1, cage 2}) in decreasing order of survival. As the
blades are placed vertically within the cabinet, pairs of nodes too
experience the same relationship with distance to the bottom, nodes 2
and 3 having lower failure rates than nodes 0 and 1.

The survival curves for the new batch are uninteresting as very few
failures have occurred in the three years they have been in operation.

We don't see a ``bathtub curve'' phenomenon, in fact the opposite is
apparent in cages 1 and 2 of the old batch. The slope of the survival
curve is related to the hazard rate. There does not seem to be an
early ``infant mortality'' period nor a ``wearout'' phenomenon at the
end.

Another standard survival data methodology is a Cox proportional
hazards (CPH) regression analys, which can include covariates and
estimate relative risk based on the covariates
\cite{Cox1972,Harrell2015}. The CPH regression function takes the form
\begin{displaymath}
  h(t) = h_0(t)e^{b_1 x_1 + b_2 x_2 \ldots b_k x_k},
\end{displaymath}
where $x_i$ are covariates, $h_0(t)$ is the {\em baseline hazard}, and
the $b_i$ are coefficients that measure the impact of the
covariates. The quantity $e^{b_i}$ is the hazard ratio for covariate
$i$.

CPH is considered a semi-parametric model as there are no assumptions
about the shape of the baseline hazard function. But its main
assumption is that the hazards are proportional. There are a number of
tests for this, including graphical diagnostics in the \pkg{survminer}
package as well as checking that survival curves for categorical
covariates do not cross. We ran these diagnostics and concluded that
the hazards for our location categories are approximately
proportional. However, the {\em batch} survival functions do cross and
so we fit the model to the new and the old batches separately. The
results are presented in Figures~\ref{fig:cox-old}
and~\ref{fig:cox-new}.
\begin{figure}
  \centering
  \includegraphics[width=0.8\columnwidth]{figs/cox_o001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt old}
    batch.}
  \label{fig:cox-old}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=0.8\columnwidth]{figs/cox_n001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt new}
    batch.}
  \label{fig:cox-new}
\end{figure}

\fix{Need to add interpretation}
